Generalized spaces for constructive algebra

### 2020-04-08 paper-herrsching2019/main.tex
Meta-introduction

\begin{abstract}
XXX abstract
\end{abstract}
\body

\tableofcontents

@@

There is only one caveat to keep in mind: Irrespective of its status in the
metatheory, the interpretation of the law of excluded middle with the sheaf
semantics is, for most locales~$X$, false (Example~\ref{ex:negneg}). People who
are familiar with other flavors of constructive mathematics should also observe
that the interpretation of any form of the axiom of choice, even countable
choice or weak countable choice, is false for most locales~$X$. (For discrete
locales, these axioms pass from the metatheory to the universe of sheaves.)

>>>>

\begin{abstract}
XXX This contribution gives an introduction to the theory of locales,
\end{abstract}
\body

\tableofcontents
\bigskip

\noindent
The purpose of this contribution is to give a coherent account of a particular
narrative which links locales, geometric theories, sheaf semantics and
constructive commutative algebra. We are hoping to convey a firm grasp of three
ideas:

\begin{enumerate}
\item Locales are a kind of space in which opens instead of points are
fundamental.
% Any propositional geometric theory gives rise to a locale of models.
\item Sheaf semantics allows us to explore mathematical objects from custom-tailored
mathematical universes.
\item Without loss of generality, any reduced ring is a field.
\end{enumerate}

Unlike some other texts on locales or sheaves, this text is set in a
constructive but impredicative metatheory. We do not use the law of excluded
middle nor any version of the axiom of choice, and we also do not adopt any
nonclassical principles. We freely use powersets, but more for linguistic
convenience than by necessity.

In particular, the central idea presented in
Section~\ref{sect:appl}, the reduction technique by which we can pretend that a
given reduced ring is a field, is very robust with respect to the metatheory.
Even though we chose a semantic point of view for its derivation, the final
result is a pure syntactical transformation. As it stands, this text could be
formalized in the kind of type theory which is valid in
toposes~\cite{maietti:modular-correspondence} or in~\textsc{izf}~\cite{crosilla:cst-izf}.

For readers familiar with the topics of this contribution, we would like to
make a couple of remarks. We present a self-contained proof of the simple
version of Barr's theorem as Theorem~\ref{thm:baby-barr}. For reducing
intuitionistic provability to geometric provability, we employ (what amounts
to) the syntactic site, similar to the refined completeness result
by Thierry Coquand~\cite{coquand:site} but closer to the standard description
of the syntactic site. For reducing classical provability to intuitionistic
provability, we employ a variant of the double negation translation. We were
not able to track down a reference to this approach in the context of
infinitary logic, but the basic idea has long been known and summarized, for
instance, in Ref.~\refcite{aczel:russell-prawitz}. % XXX first-order completion. We probably should introduce sites but don't

\subsection*{Introduction}
\addcontentsline{toc}{section}{Introduction}

@@

There is only one caveat to keep in mind: Irrespective of its status in the
metatheory, the interpretation of the law of excluded middle with the sheaf
semantics is, for most locales~$X$, false (Example~\ref{ex:negneg}).
Practitioners of other flavors of constructive mathematics should also observe
that the interpretation of any form of the axiom of choice, even countable
choice or weak countable choice, is false for most locales~$X$. (About the only
exception are locales induced from discrete topological spaces. For such
locales, the law of excluded middle and countable, dependent or full choice
pass from the metatheory to the universe of sheaves.)
